### 2010-01-14 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
First initial working draft of D2.4.2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@419 d8b0c59b-6544-0410-94d9-83a5658a058e

\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{textcomp} %vertical apostrophe
\usepackage{listings} %source code listings
\usepackage{amssymb} %math symbols
\usepackage{graphicx} %image placement
\usepackage{supertabular}
\parindent0pt

\newcommand\deliverableNumber{D2.4.1}
\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST ontology}
\newcommand\deliverableTitleShort{Mediation amongst ontologies}
\newcommand\workpackageNumber{2}
\newcommand\workpackageTitle{Definition of Conceptual Model}
\newcommand\authorOne{Oszkar Ambrus, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

@@

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 27/02/2009}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
							\textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 27.02.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
\end{tabular}
\end{small}

@@

We review current work and literature, discussing the main areas of the
topic, that is the core problem of ontology mediation, the types of
mismatches that occur between ontologies, and the three main ontology
mediation approaches, namely (1) the representation of correspondences
between ontologies, (2) the (semi-)automatic process of discovering
these correspondences and (3) creation of a new ontology from the
source ontologies.

By the overview of current research in the field of ontology mediation,
we lay the basis of the implementation of ontology mediation in FAST.

@@

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.1 is the first iteration of the FAST ontology mediation approach. This document contains a general overview of current literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation. Based on this general overview, the implementation of ontology mediation within FAST can be researched, which will constitute the topic of the next iteration of this document.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment, ontology merging}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:introduction}
\paragraph{Goal and Scope.}
The central driver for Enterprise Mashups is the lightweight resource
composition style to build enterprise applications with reusable
building blocks, called Gadgets. The architecture used for this
front-end layer of Gadgets specifies a back-end layer of Services,
designed to support machine-to-machine interaction over a network (e.g.
Web Services, Databases, etc.). Having this case, the communication
problems need to be addressed, that arise from disparate interfaces,
different invocation mechanisms and even different conceptualisations
of the data transmitted between the front-end and the back-end. This
document deals with the third issue, namely the differences between the
conceptualisations of the exchanged data.

As in all fields employing semantic web paradigms, data is annotated
using ontologies. Ontologies conceptualise the data used, describe
their context and their semantic background that should be known to all
agents using that data. Because we do not the have the situation where
the same ontologies are used by all sides, having different ontologies
and even different languages to describe them, we need to integrate the
ontologies present in the back-end with the ones present in the
front-end, and to find a way to mediate between the the different
ontologies used in the two layers.

We try to find a solution to these differences which range from simple
naming differences to more complex data heterogeneity, behavioural
differences, discrepant paradigms and other disparities. The
reconciliation of these differences is called ontology mediation (often
mentioned in the specialised literature as ontology matching, or
ontology alignment, thus we will sometimes use these terms as well). 
\paragraph{Structure of the Document.}
We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology
mediation problem in Section \ref{sec:mediationproblem} , discuss the types of mismatches between
ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology
mediation in Section \ref{sec:mediationapproaches}.
\paragraph{Overall Approach for WPX and relation to other Work Packages and Deliverables.}
Ontology mediation will play an important role for the integration of 
external ontologies into the FAST ontology infrastructure.
By giving a detailed overview of the state-of-the-art in ontology
mediation, we lay the groundwork for further research in T2.4 [Task
2.4, see Project Description] with this first iteration of D2.4. Future
versions of the deliverable (i.e., D2.4.2 and D2.4.3) will implement
ontology mediation within the FAST platform.

\section{Ontologies}
\label{sec:ontologies}
This section provides a brief introduction to the concept of ontologies.

\subsection{Definitions}
As specified in \cite{gruber2008ontology} {\textquotedblleft}an ontology defines a
set of representational primitives with which to model a domain of
knowledge or discourse{\textquotedblright}. These primitives are
typically classes, attributes and relations among classes or class
instances (and others, such as function terms, rules, restrictions,
axioms and events). Their definitions provide information about their
semantics (meaning) and constraints on their logically consistent
usage.

Ontologies are typically specified in languages that allow to abstract
away from implementation details, and focus on the so-called semantic
level. This way ontologies provide a representation independent from
data models and implementation, thus providing a means to enable
interoperability among disparate systems and specifying interfaces to
independent, knowledge-based systems. In the technology stack of the
Semantic Web, ontologies define a separate layer, and have a set of
languages and tools (both commercial and open-source) to work with
them.

We would also mention a widely-known and used classical definition of
ontologies, presented in \cite{gruber93towards}: {\textquotedblright}[An
ontology is] an explicit, formal specification of a shared
conceptualisation{\textquotedblright}. \cite{studer1998knowledge_engineering_principles} explains
this definition, by elaborating on the four main terms of the
definition:

\begin{enumerate}
\item it is a conceptualisation, because it models and categorises
relevant concepts from the real world, 
\item it is explicit, because the model explicitly states the types of
the concepts, the relationships between them and the constraints of
their use, 
\item it is formal, because the ontology has to be machine-readable and
consistent, and 
\item it is shared, because an ontology is consensual, i.e. accepted by
a group of people and it is used in a shared environment to allow a
common conceptualisation of data. 
\end{enumerate}

\subsection{General Description}
An ontology is used to define the vocabulary that agents use to exchange
queries and assertions. If an agent commits to a common ontology, it
guarantees the consistency (but not completeness) of the queries and
assertions using the vocabulary defined in the ontology \cite{gruber93towards}.
An agent supporting the interface defined in an ontology is not
required to use the terms of the ontology as an internal encoding of
its knowledge. Nonetheless, the definitions and formal constraints of
the ontology do restrict what can be meaningfully stated in the
language defined.

Two analogies to ontologies are given in \cite{gruber93towards}. Ontologies are
similar to global type declarations in a conventional software library,
and ontological commitments are similar to type restrictions over the
inputs and outputs of program modules. Formal argument restrictions can
be checked mechanically by compilers, to make sure that calling
procedures pass legal data to called procedures. Similarly, sentences
in an exchange between agents can be checked for logical consistency
using the definitions in the ontologies. Just as the formal argument
list hides the internal structure of a procedure from its calling
environment, a common ontology allows one to interact with a
knowledge-based program without committing to or being aware of its
internal encoding of knowledge.

Ontologies are also like conceptual schemata in database systems. Such a
schema provides a logical description of the data being shared,
allowing applications to interact with it without having any knowledge
of internal data structures. So while a conceptual schema defines
relations and constraints on data, an ontology defines terms with which
to represent knowledge, and constraints on their relationships.
\cite{gruber2008ontology} also points out that the requirement of inputs and
outputs to be logically consistent with the definitions and constraints
of an ontology is analogous to the requirement that the rows of a
database table (or insert statements in SQL) must be consistent with
the integrity constraints imposed on the table, stated separately and
independently of the internal data formats.

\cite{gruber2008ontology} enumerates the key applications of ontologies, by saying
that they {\textquotedblleft}are part of the W3C standards stack for
the Semantic Web, in which they are used to specify standard conceptual
vocabularies in which to exchange data among systems, provide services
for answering queries, publish reusable knowledge bases, and offer
services to facilitate interoperability across multiple, heterogeneous
systems and databases.{\textquotedblright} The key role of ontologies
in database systems is to provide a data modeling layer above specified
database designs, so that data can be used across independently
developed systems and services. This made it possible to achieve
database interoperability, cross database search and web service
interrogation.

There are many issues not addressed by common ontologies, related to
knowledge sharing. One important question is how a group of people can
reach a consensus on a common ontology. The discussion of ontology
mediation aims to address this problem by attempting a solution to the
case of several ontologies conceptualising a similar domain of the real
world in different ways.

\section{The Mediation Problem} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:mediationproblem}
Ontologies are employed as explicit descriptions of the information
source semantics. According to \cite{wache2001ontology_information_integration}, integrating these
semantics along multiple systems can be done using three main
approaches. (1) The single ontology approach, using a single ontology
as a global shared vocabulary. Unfortunately this can not provide
different views on a domain, and using it for multiple similar
catalogues is unfeasible due to the vast amount of information needed
to be captured, and the susceptibility of the conceptualisation to
change. (2) The multiple ontology approach has each information source
describing its own ontology. This solves problems related to the single
ontology approach, but introduces the problem of differences in
conceptualising the same concepts. Ontology mediation can be used to
map different parts of the ontologies to each other. (3) The hybrid
approach uses multiple ontologies, but uses a global ontology to make
all the ontologies comparable to each other. This is closely related to
ontology merging, where overlapping entities are merged, and the rest
is simply composed into one unified ontology, allowing for the
definition of the concepts of both ontologies. We suggest that the FAST
Gadget Ontology (as defined in D2.1) could act as a global ontology in
the context of FAST, although further research needs to be done in this
direction, since the Gadget Ontology has a slightly different
epistemological commitment.

In the context of FAST, the problem is that of reconciling the
ontologies in the back-end with the ontologies in the front-end, to
allow for interoperability between the two layers.

In the following we present a motivating scenario for ontology mediation
in Section \ref{subsec:mediationproblem_motivatingexample}, and a formal description of the alignment problem and
process, respectively, in Section \ref{subsec:mediationproblem_problemstatement}. After that we show some examples
of ontology mediation in Section \ref{subsec:mediationproblem_examples}.

\subsection{Motivating Example}
\label{subsec:mediationproblem_motivatingexample}
This section presents a brief example, illustrating a problem that
presents the need for ontology mediation.

%\begin{verbatim}
\lstset{frame=single, basicstyle=\ttfamily, caption=Ontology \texttt{O1}.}
\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
\lstset{emphstyle=\textbf\textit\underbar}
\begin{lstlisting}
vsmlVariant -"http://www.smo.org/wsml/wsml-syntax/wsml-flight"

namespace { -"http://see.deri.org/adrian/thesis/ontologies/O1#" ,
       wsml - "http://www.wsmo.org/wsml/wsml-syntax#" ,
         dc - "http://purl.org/dc/elements/1.1/" }

ontology - "http://see.deri.org/thesis/ontologies/O1"
    nonFunctionalProperties
    dc#description hasValue "A simple ontology modeling
                             the concept of a person"
    endNonFunctionalProperties

concept person
    name ofType _string
    age ofType _integer
    hasGender ofType gender
    hasChild ofType person
    marriedTo ofType person

concept gender
    value ofType _string

instance male memberOf gender
    value hasValue "male"
instance female memberOf gender
    value hasValue "female"
\end{lstlisting}

Consider the two ontologies \texttt{O1} and
\texttt{O2} \cite{mocan2008mediation} presented in Listing 1 and
Listing 2 respectively, in the Human Readable Syntax of the Web Service
Modeling Language (WSML) \cite{debruin2005wsml}. Elements from
ontology \texttt{O1} need to be mapped to the elements
described in Ontology \texttt{O2}.

Ontology \texttt{O1} describes the concept
\texttt{person} as one having five attributes, each of
them having a type that is either a concept or a value from a given
data type. The concept gender has two instances defined, that have
attributes pointing to the corresponding values.

\lstset{frame=single, caption=Ontology \texttt{O2}.}
%\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
%\lstset{emphstyle=\textbf}
\begin{lstlisting}
vsmlVariant -"http://www.smo.org/wsml/wsml-syntax/wsml-flight"

namespace { -"http://see.deri.org/adrian/thesis/ontologies/O2#" ,
       wsml - "http://www.wsmo.org/wsml/wsml-syntax#" ,
         dc - "http://purl.org/dc/elements/1.1/" }

ontology - "http://see.deri.org/thesis/ontologies/O2"
    nonFunctionalProperties
    dc#description hasValue "A simple ontology modeling
                             the concept of a human"
    endNonFunctionalProperties

concept human
    name ofType _string
    age ofType _integer
    noOfChildren ofType _integer

concept man subConceptOf human

concept woman subConceptOf human

concept marriage
    hasParticipant ofType human
    date ofType _date
\end{lstlisting}

%\newline
Ontology \texttt{O2} describes the concept
\texttt{human} as one having three attributes. The
concepts \texttt{man} and \texttt{woman} are
subclasses of the concept \texttt{human}.

Additionally a fourth concept called \texttt{marriage} is
defined having two attributes. (For simplicity we do not deal with the
exceptions of polygamy and same-sex marriage).

Suppose that a service (or agent) using ontology O1 needs to interact
with a service accepting and providing data that is semantically
defined in ontology O2.

The first step is to identify candidates to be mapped or to have
taxonomic relationships under an integrated schema \cite{shvaiko2005schema_based}. It is obvious that \texttt{person} and
\texttt{human} are candidates to be mapped, as well as the
class \texttt{man} with the class
\texttt{person} having the attribute
\texttt{gender} equal to
{\textquotedblleft}\texttt{male}{\textquotedblright} and
the class \texttt{woman} with the class
\texttt{person} having the attribute
\texttt{gender} equal to
{\textquotedblleft}\texttt{female}{\textquotedblright}.

After having identified these correspondences, the next step is to
generate query expressions and assertions that automatically translate
data instances of one of these ontologies into the other one, or that
automatically handle their semantic relations under an integrated
ontology.

\subsection{Problem Statement}
\label{subsec:mediationproblem_problemstatement}
Given two ontologies \texttt{o} and
\texttt{o{\textquotesingle}} that need to be mapped to
each other, according to \cite{shvaiko2005schema_based} we can define a
mapping element as a 5-tuple \texttt{{\textless}id, e,
e{\textquotesingle}, n, R{\textgreater}}, where

\begin{itemize}
\item \texttt{id} is a unique identifier, identifying the
mapping element, 
\item \texttt{e} and
\texttt{e{\textquotesingle}} entities (classes,
attributes, relations) of the first and second ontology, respectively, 
\item \texttt{n} is a confidence measure in some
mathematical structure holding the correspondence value between e and
e{\textquotesingle}, 
\item \texttt{R} is the correspondence relation holding
between e and e{\textquotesingle} (e.g.,
\texttt{equivalence (=); more general($\sqsupseteq$); disjointness($\perp$); overlapping($\sqcap$)}.
\end{itemize}

\begin{figure}
    \centering
        \includegraphics[width=72mm, height=45mm]{images/Alignment_process.png}%[bb=0 0 201 126]
        \caption{The alignment process}
    \label{fig:Alignment}
\end{figure}

A mapping is a set of the above defined mapping elements. The alignment
operation determines the mapping
(M\texttt{{\textquotesingle}}) for the pair of ontologies
(\texttt{o} and
\texttt{o{\textquotesingle}}). There are some parameters
than can extend the definition of the alignment process, namely:

\begin{enumerate}
\item an input mapping, \texttt{M}, which is to be used and
completed by the process, 
\item the mapping parameters, \texttt{p} (such as weights
and thresholds), that configure the alignment process, 
\item other external resources used by the alignment,
\texttt{r} (such as dictionaries, thesauri, etc.); See Figure \ref{fig:Alignment}. 
\end{enumerate}
For example, suppose that based on some algorithm (using linguistic and
structural analysis) the confidence measure for the equivalence
relation between the classes \texttt{person} in
\texttt{O1} and \texttt{human} in
\texttt{O2} is 0.62. Suppose that a mapping parameter
specifies a threshold of 0.55 for this algorithm, that is all pairs of
entities with a confidence measure higher that 0.55 will be considered
correct mapping elements. Thus our algorithms would return the
following mapping element: \texttt{{\textless} id11,
person, human, 0.62, = {\textgreater}.}Now, another algorithm
determines the two concepts to mean exactly the same thing, thus it
either does not compute a confidence measure (returning
\texttt{n/a}), or it returns the upper limit of the
confidence measure interval (e.g. 1). Thus the mapping element would
be: \texttt{{\textless} id11, person, human, n/a, =
{\textgreater}.}

It should be noted, that the entities mapped between two ontologies can
be constructed using patterns, i.e. more than just mapping concepts to
concepts, and so on. For example, we can map a concept to a concept
having a certain attribute, like in the case of the concept
\texttt{person} having the attribute
\texttt{hasGender} equal to \texttt{female}
from \texttt{O1} being mapped to the concept
\texttt{woman} from \texttt{O2}.

\subsection{Mediation Examples}
\label{subsec:mediationproblem_examples}
We present two further examples to better illustrate the use of ontology
mediation.

The first example is taken from \cite{giunchiglia2008evaluation}. Based on the
idea that we view ontologies as graph-like structures, we exemplify
ontology matching with the help of fragments of two tree-like
structures, such as Google and Looksmart. Notice that in the general
case the relation holding between nodes is not specialisation, but
classification (parent-child relation).

\begin{figure}[h]
    \centering
        \includegraphics[width=101mm, height=67mm]{images/F2GoogleLooksmart.png}%[bb=0 0 254 167]
        \caption{Fragments of Google and Looksmart directories and some
correspondences. Double arrowheads denote equivalence, while single
arrowheads stand for the more general relation. Image taken from
\cite{giunchiglia2008evaluation}.}
    \label{fig:GoogleLooksmart}
\end{figure}

Suppose we want to merge the merge the two structures. Such situations
can arise, for example when an e-commerce company acquires another one.
One probable mapping is given in Figure \ref{fig:GoogleLooksmart}. We have found, for example,
that \texttt{O1:Basketball} is equivalent to
\texttt{O2:Basketball}, and that
\texttt{O2:Games} is more general than
\texttt{O1:Board\_Games.}

The second example given in \cite{madhavan2001schema_matching} illustrates the
need to use several techniques in matching two ontologies. Figure \ref{fig:POSchemas}
shows two purchase order schemas that need to be matched.

\begin{figure}[h]
    \centering
        \includegraphics [width=101mm, height=42mm]{images/F1POSchemas.png}%[bb=0 0 257 105]
        \caption{Matching purchase order schemas. Image taken from \cite{madhavan2001schema_matching}.}
    \label{fig:POSchemas}
\end{figure}

As stated above in Section 3.2 we look for similarity coefficients
between the two schemas and deduce a mapping from those coefficients.
Suppose, as a first phase we use linguistic matching, using thesauri to
help match names, acronyms, synonyms, etc. resulting in a set of
candidate pairs (such as \texttt{PO} and
\texttt{PurchaseOrder}, or \texttt{Bill} and
\texttt{Invoice}) and language similarity coefficients
(\texttt{lsim}). In the second phase we would perform a
structural matching of the two schemas. For example,
\texttt{Line} would be mapped to
\texttt{ItemNumber}, because their parents and their other
two children match, \texttt{City} under
\texttt{POBillTo} will be mapped to
\texttt{City} under \texttt{InvoiceTo},
because bill is a synonym to invoice. This process results in
structural similarity coefficients (\texttt{ssim}). By
using a weighted similarity, we can compute the final mappings (e.g. by
using the formula \texttt{wsim=ws*wsim+(1-ws)*lsim}, where
\texttt{ws} is the interval [0,1]).

\section{Ontology Mismatches}
\label{sec:ontologymismatches}
Ontology mismatches are the main obstacles in the combined use of
independently developed ontologies. We will now give an overview of the
main type of ontology mismatches as presented in \cite{klein2001combiningOntologies}, which
is the main source of input to this state-of-the-art report. We will
classify ontology mismatches into given types and will find their
relationships.

Firstly, two levels of mismatches can be distinguished. The first level
is the \textit{language} or meta-model level. These are differences in
the primitives that make up the language used to specify the ontology.
They produce mismatches in the mechanics to define classes, relations,
attributes and so on. The second level is the \textit{ontology} or
model level. This is were the actual ontology is located. These are
differences in the way the domain is modeled.

In this section we discuss the overlap and mismatches between concepts,
relations, attributes and instances. We will detail the two levels of
mismatches and the different types that occur at each level.

\subsection{Language Level Mismatches}
\label{subsec:ontologymismatches_languagelevel}
We define four types of mismatches, although they often coincide. These
occur when ontologies written in different languages are combined. We
note that these type of mismatches will be less relevant for FAST
because we assume that standards for ontology languages are now so well
established that in practice we will not very often find ontologies
written in different languages anymore, but we discuss them briefly to
give a better overview of the meditation problem.

\subsubsection{Syntax}
Different languages obviously use different syntax. Therefore the same
concept can (and will be) often modeled in a completely different way.
For example, the concept of chairs would be defined in RDF Schema as
\texttt{{\textless}rdfs:Class
ID={\textquotedblright}Chair{\textquotedblright}{\textgreater}}, while
in LOOM it would be \texttt{(defconcept Chair)}.

This is the simplest possible mismatch, although it often comes coupled
with other mismatches as well. The simplest form is when a language has
several syntactical forms.

The solution is to simply to have a rewrite mechanism that returns the
corresponding translated form.

\subsubsection{Logical representation}
A slightly more complicated mismatch is when a logical notion can not be
represented in a given language. For example, in certain languages it
is possible to assert that two classes are disjoint, say using
\texttt{(disjoint A B),} whereas in other languages it is
necessary to express this using simpler statements, such as the
negation of set-inclusion statements, like \texttt{A
subclass-of (NOT B), B subclass-of (NOT A).}

The actual problem is not the question if something can be expressed,
but finding the language constructs to express a logical notion (not a
concept).

Solving this mismatch is relatively-easy: it consists in finding
translation rules to transform one representation into another.

\subsubsection{Semantics of primitives}
This difference is a more subtle one, and occurs at the meta-model of
the language. It sometimes occurs that the semantics of an identical
name differs from language to language. For example, the expression
\texttt{A equalsTo B} can have many different
interpretations.

Even when two ontologies use the same syntax, the semantics may differ.
For example, the OIL RDF syntax interprets several
\texttt{{\textless}rdfs:domain{\textgreater}} statements
as the intersection of the arguments, whereas RDF Schema interprets
them as the union of the arguments.

\subsubsection{Language expressiveness}
This mismatch, at the meta-model level of the languages, is the one
having the greatest impact on the attempt to combine and reconcile
ontologies.

The difference implies that some languages can express notions that
others can not. For example, some languages have no constructs to
express negation, whereas other do. Other typical examples are support
for sets, lists, etc.

These type of mismatches have the greatest impact, and are sometimes
mentioned as the {\textquotedblleft}fundamental
differences{\textquotedblright} between knowledge models.

\subsection{Ontology Level Mismatches}
\label{subsec:ontologymismatches_ontologylevel}
Mismatches at the model level happen both in the case of having the two
ontologies written in the same language and syntax and in the case of
different languages. We can distinguish between several types of
ontology level mismatches.

A conceptualisation mismatch is a difference in how the domain is
modeled (conceptualised), therefore we will have different concepts and
different relations between them. Explication mismatches are
differences in the way the concepts are specified. These occur as
mismatches in definitions, or terms, or the combination of the two.
Terminological mismatches are naming differences or conflicts. Finally,
encoding mismatches are differences in the way values are specified,
for example currencies for prices.

>>>>

\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}        % turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{lipsum}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{textcomp} %vertical apostrophe
\usepackage{listings} %source code listings
\usepackage{amssymb} %math symbols
\usepackage{graphicx} %image placement
\usepackage{supertabular}

\usepackage{amsmath}
\usepackage{pgf} %graph plot
\usepackage{tikz}
\usepackage{paralist}
\usetikzlibrary{plotmarks}

\parindent0pt

\newcommand\deliverableNumber{D2.4.2}
\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST platform}
\newcommand\deliverableTitleShort{Mediation amongst ontologies}
\newcommand\workpackageNumber{2}
\newcommand\workpackageTitle{Definition of Conceptual Model}
\newcommand\authorOne{Oszkar Ambrus, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

@@

%\begin{minipage}[p]{15cm}
    \begin{center}
        \includegraphics{images/FAST_logo}\\
        \vspace{1cm}
        {\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
        \vspace{4cm}
        {\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \textbf{\deliverableTitle}}\\
        \vspace{2cm}
        {\large \sffamily \authorOne}\\
        \vspace{0.5cm}
        \vfill
        {\large \sffamily Date: 20/02/2010}\\
        \vspace{1cm}
        {\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
                            \textcolor{white}{\textbf{Change description}}\\ \hline
0.1 & 07.01.2010 & Oszkar Ambrus (NUIG) & initial draft \\ \hline
%0.2 & 19.02.2009 & Oszkar Ambrus (NUIG) & content moved from wiki after internal review \\ \hline
%0.1 & 07.01.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
\end{tabular}
\end{small}

@@

In D2.4.1 we reviewed current work and literature, discussing the main areas of the topic, that is the core problem of ontology mediation, the types of mismatches that occur between ontologies, and the three main ontology mediation approaches, namely (1) the representation of correspondences between ontologies, (2) the (semi-)automatic process of discovering these correspondences and (3) creation of a new ontology from the source ontologies.

Based on the theoretical foundations laid down in D2.4.1 we present our approach to integrating ontology matching into the FAST platform. We present the roles related to ontology matching in FAST, and the way ontology matching is achieved within each one of those roles.

@@

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.2 is the second iteration of the FAST ontology matching approach. The firsty iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, this current interation presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:introduction}
\paragraph{Goal and Scope.}
FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones. Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets''. In the services created within the FAST environment, the ontologies come from different parties involved in the creation of gadgets and building blocks. These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary, as the means to reconcile differences in the various conceptualisations.

We aim to integrate ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.  We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities. The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).

\paragraph{Structure of the Document.}
Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment. 
%Describe the rest.
Part \ref{part:d2.4.1} contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching. We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology mediation problem in Section \ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in Section \ref{sec:mediationapproaches}.

\paragraph{Overall Approach for WPX and relation to other Work Packages and Deliverables.}
Ontology matching will play an important role for the integration of different underlying resources in the user-built gadgets. 
We present our initial research in the implementation of ontology matching in this second iteration of D2.4, after having laid the theoretical ground work of ontology matching in the first iteration. The final version of the deliverable (i.e., D2.4.3) will finalize the implementation of ontology mediation within the FAST platform.

\part{Implementing Ontology Matching within the FAST Platform}
\label{part:d2.4.2}

\include{fast_d2_4_2_part1}

\part{D2.4.1: Theoretical Foundations for Ontology Mediation}
\label{part:d2.4.1}
%\include{fast_d2_4_2_part2}

\section{Summary}
\label{sec:summary}
With the Semantic Web winning more and more ground both in research and industry, the mismatches between ontologies used by the inter-operating systems present an increasing problem. Ontology matching deals with identifying these mismatches, and overcoming them.

In the first iteration (described in Part \ref{part:d2.4.1} we laid the theoretical foundations of ontology matching. After presenting the main problem, and the types of mismatches between ontologies, we have discussed - based on the most prominent research results in these areas - the three main approaches to ontology mediation, that is \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging }(the process of creating a new ontology based
on the source ontologies that need to be reconciled).

In this second iteration we established the approach to integrate ontology matching into FAST, and have shown the implementation details as well as test scenarios. We have presented a tool for finding the alignments between ontologies, as well as a powerful language to represent them, beyond the capabilities of the tool. We have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.

We plan to evaluate the feasibility of ontology matching in real-life tests and finalize it for the completed version of the FAST platform.
